---
title: "Reproducible Research"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(dplyr)
require(ggplot2)
```

## Course Project 1

### What is mean total number of steps taken per day?
1. The total number of steps taken per day.
```{r }
activity<-read.csv(file = "/Users/alexeypogorelov/Downloads/activity.csv")
activity<-tbl_df(activity)
output1<-activity %>%
    group_by(date) %>%
      summarise(tot_step=sum(steps))
```
2. Histogram of the total number of steps taken each day

```{r}
ggplot(output1,aes(tot_step))+geom_histogram()+ggtitle("Histogram of total steps taken per day")+xlab("")+ylab("")

```

3. The mean of the total number of steps taken per day is 
`r format(mean(output1$tot_step,na.rm = TRUE))`, while the median is `r median(output1$tot_step,na.rm = TRUE)`

### What is the average daily activity pattern?

1. The average number of steps taken, averaged across all days (y-axis)

```{r}
output2<-activity %>%
  group_by(interval) %>%
  summarise(av_step=mean(steps,na.rm = TRUE))

ggplot(output2,aes(x=interval,y=av_step))+geom_line()+ggtitle("Average number of steps taken")+xlab("5-minute interval")+ylab("")

```

2. Interval number `r as.numeric(arrange(output2,desc(av_step))[1,1])` on average across all the days in the dataset contains the maximum number of steps.

### Imputing missing values.

1. The total number of missing values in the dataset `r sum(is.na(activity))`

2. Let's substitute missing values with the mean of 5-min interval
```{r}
activity3<-full_join(activity,output2,by="interval")

activity3$steps<-ifelse(is.na(activity3$steps),activity3$av_step,activity3$steps)
activity3$av_step<-NULL
```

3. The new data set where missing values substituted with interval means 
```{r}
head(activity3)
```

4.

```{r}
output3<-activity3 %>%
  group_by(date) %>%
  summarise(tot_step=sum(steps))

ggplot(output3,aes(tot_step))+geom_histogram()+ggtitle("Histogram of total steps taken per day")+xlab("")+ylab("")

```

5.The mean of the total number of steps taken per day is 
`r format(mean(output3$tot_step))`, while the median is `r format(median(output3$tot_step))`

### Are there differences in activity patterns between weekdays and weekends?


1. 

```{r}
activity2<-activity3 %>%
  mutate(factor=ifelse(weekdays(as.Date(activity3$date),abbreviate=TRUE) %in% c("Sat","Sun"),"weekend","weekday"))

```

2.

```{r }
output4<-activity2 %>%
  group_by(interval) %>%
  filter(factor=="weekday") %>%
  summarise(av_step=mean(steps,na.rm = TRUE))

ggplot(output4,aes(x=interval,y=av_step))+geom_line()+ggtitle("Average number of steps taken during weekdays")+xlab("5-minute interval")+ylab("")

```


```{r}
output5<-activity2 %>%
  group_by(interval) %>%
  filter(factor=="weekend") %>%
  summarise(av_step=mean(steps,na.rm = TRUE))

ggplot(output5,aes(x=interval,y=av_step))+geom_line()+ggtitle("Average number of steps taken during weekends")+xlab("5-minute interval")+ylab("")

```

